---
title: "Penguins_anova"
author: "Gracie White"
date: "November 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. load packages and read in the penguins data

```{r, include=FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read.csv("penguins.csv")

```
Our question is, is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)? 

2. do some exploring. Note that data is already in tidy format! yay!

New functions:
with + table for easy counting over multiple factors 

Update the species Name using case_when 


```{r}
counts <- with(penguins, table(species, sex))

penguins2 <- penguins %>% 
  mutate(common_name = 
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
           )) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")


```


- Exploratory Graphs
- Summary Statistics 
- Levene's test for equal variances 
- 1-way ANOVA
- Post-hoc testing!


```{r}
hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~common_name)

hists
```


```{r}
qqplots <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq() + 
  facet_wrap(~ common_name)

qqplots
```

```{r}
# now Lavene's test for equal variances, (> 2 groups, use the F test)

#we will use the function leveneTest from the 'car' package
#H0: variances are equal 

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)
penguin_levene

# We retain the null hypothesis of equal variances 

var_table <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

var_table
#highest var is still less than 4x the smallest variance 


```

Part 3: run a one-way ANOVA

Single Factor: common_name (species)
Number of levels in the factor: 3 (chinstrap, gentoo, adelie)
random variable: flipper_length (mm)

Is there a sig diff in mean flipper length for female penguins by species? 

```{r}
penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

#H0: there is no difference in means between groups, mean flipper lengths across all species are equal
#HA: there is at least one signifcant differance in means 


#we can conclude that at least two samples were taken from samples with different means 
#there is a significant difference in flipper lengths based on species (F(df) = statistic, p < 0.001, $alpha\$ = anova 

#so now we need to figure out which ones are different 
#post-hoc testing
penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

#mean flipper lengths are all significantly different with post-hoc Tukey's HSD, alpha = 0.05



```

we can conclude that at least two samples were taken from samples with different means 
there is a significant difference in flipper lengths based on species (F(df) = statistic, p < 0.001, $alpha\$ = anova 


